---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Insert title of project here"
subtitle: "Web address for GitHub repository"
author: "Name"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

# Rationale and Research Questions
> Our main research question was if NC counties with higher concentration of hog farms would have higher health disparities compared to counties with lower hog farm concentrations. To research this, we looked at health and population distribution data for each NC county and specifically looked at data for counties with the highest concentrations of hog farms (Bladen, Duplin, Greene, Sampson, Robeson, Wayne) and comparing them to Durham as we are current residents of this county. We wanted to gain a better understanding of the inequities and environmental justice issues that stem from NC hog farms and pork production economy. 

\newpage

# Dataset Information
>Our dataset is...........

\newpage

```{r setup, include=FALSE}
# Set your working directory
getwd()

# Load your packages
#install.packages("tidyverse")
library(tidyverse)
#install.packages("magrittr")
library(magrittr)
#install.packages("dplyr")
library(dplyr)
#install.packages("sf")
library(sf)



# Set your ggplot theme
mytheme <- theme_classic(base_size = 12) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

# Load your datasets
Health_Data_Raw <- read.csv("../ChanJacob_ENV872_EDA_FinalProject/Data/Raw/County Health Data 2021.csv", stringsAsFactors = TRUE)

```

# Exploratory Analysis 

##Data Wrangling

```{r, data wrangling}
##Wrangling the dataset 

#Removing columns and rows containing weblinks and other metadata 
Health_Data_Clean <- Health_Data_Raw[-c(1,2,3),-c(1,3, 4, 5, 6)] 

#Using a pipe to select and rename columns
Health_Data_Clean <- Health_Data_Clean %>%
  select(Bold.Indicator.Name.Text, NC, X.6, X.8, X.30,X.31, X.38, X.39, X.57, X.81, X.77, X.95, X.97) %>%
  rename(Indicators = Bold.Indicator.Name.Text, Bladen = X.8, Duplin = X.30, Durham = X.31, Greene = X.39, Sampson = X.81, Robeson = X.77, Wayne = X.95, Wilson = X.97, Granville = X.38, Martin = X.57, Beaufort = X.6)
#Durham - 

#Transposing the data
Health_data_transposed <- t(Health_Data_Clean)

#Converting the transposed data to a dataframe
Health_data_transposed <- as.data.frame(Health_data_transposed)

#Adding a pipe statement to rename the columns
Health_data_transposed_clean <- Health_data_transposed %>%
  rename(Population = 1, Child_Population_percentage = 2, Elderly_Population_percentage = 3, White_Population_percentage = 4, Hispanic_Latinx_percentage = 5, African_American_percentage = 6, American_Indian_percentage = 9, Reading_Proficiency_percentage = 12, Uninsured_Adults_percentage = 14, Medicaid_CHIP_enrolles_percentage = 16, Primary_Care_Physicians = 17, Low_Birthweight_percentage = 32, Infant_Mortality_Rate = 33, Poverty_percentage = 41, Air_Pollution = 52) 

#Selecting the required rows 
Health_data_transposed_clean_1 <- Health_data_transposed_clean[-c(1), -c(7,8, 10, 11, 13, 15, 18:31, 34:40, 42:51, 53:59)]

#Removing '%' character from the dataframe using the 'gsub' function
Health_data_transposed_clean_1 <- Health_data_transposed_clean_1 %>%
  mutate(Population = as.numeric(gsub(",","",Population)),
         Child_Population_percentage = as.numeric(gsub("%", "", Child_Population_percentage)),
         Elderly_Population_percentage = as.numeric(gsub("%", "", Elderly_Population_percentage)),
         White_Population_percentage = as.numeric(gsub("%", "", White_Population_percentage)),
         Hispanic_Latinx_percentage = as.numeric(gsub("%", "", Hispanic_Latinx_percentage)),
         African_American_percentage = as.numeric(gsub("%", "", African_American_percentage)),
         American_Indian_percentage = as.numeric(gsub("%", "", American_Indian_percentage)),
         Reading_Proficiency_percentage = as.numeric(gsub("%", "", Reading_Proficiency_percentage)),
         Uninsured_Adults_percentage = as.numeric(gsub("%", "", Uninsured_Adults_percentage)),
         Medicaid_CHIP_enrolles_percentage = as.numeric(gsub("%", "", Medicaid_CHIP_enrolles_percentage)),
         Low_Birthweight_percentage = as.numeric(gsub("%", "", Low_Birthweight_percentage)),
         Poverty_percentage = as.numeric(gsub("%", "", Poverty_percentage)))



#Checking the class of Health_data_transposed_clean_1
class(Health_data_transposed_clean_1)

#saving wrangled health data in Processed Data folder
write.csv(Health_data_transposed_clean_1, file = "./Data/Processed/CountyHogs_HealthData.csv")
```

Analysis: 

```{r}
mapviewOptions(fgb = FALSE)


#Filtering for NC counties from US counties shapefile
nc_county_sf <- read_sf("./Data/Spatial/cb_2018_us_county_20m.shp") %>%
  filter(STATEFP == 37) #filter for NC counties

#Loading in the shapefile
NorthCarolina_sf <- st_read('../ChanJacob_ENV872_EDA_FinalProject/Data/Spatial/cb_2018_us_county_20m.shp') %>%
  filter(STATEFP == 37)


NorthCarolina_counties_sf <- NorthCarolina_sf %>%
  filter(NAME %in% c("Robeson", "Sampson", "Duplin", "Bladen", "Greene", "Wayne", "Durham"))

mapview(NorthCarolina_counties_sf)

#Loading in the Hog Farm NC locations file 

HogFarm_Locations <- read.csv("../ChanJacob_ENV872_EDA_FinalProject/Data/Spatial/Hog_Farm_Locations_NC.csv") 

#Selecting the required columns with Lat/Long
HogFarm_Locations <- HogFarm_Locations %>%
  select(title, latitude, longitude) %>%
   na.omit(HogFarm_Locations)
  
#Converting the dataframe to a sf (spatial features) dataframe
HogFarm_Locations_sf <- HogFarm_Locations %>%
  st_as_sf(coords = c('longitude', 'latitude'),
           crs = 4269)

colnames(HogFarm_Locations_sf)


HogFarm_Locations_plot <-
ggplot() +
  geom_sf(data = nc_county_sf, aes(fill = 'turquoise')) +
  geom_sf(data = HogFarm_Locations_sf, aes(color = 'red'), size = 1) + 
  scale_fill_identity(name = "", breaks = c("turquoise"), 
                       labels = c("North Carolina"), guide = "legend") +
  scale_color_identity(name = "", breaks = c('red'), 
                       labels = c('Hog Farms'), guide = 'legend') +
  ggtitle("Hog Farms in North Carolina") +
  mytheme 
print(HogFarm_Locations_plot)


Intersection <- HogFarm_Locations_sf[NorthCarolina_counties_sf,]

HG <- 
  ggplot() +
   geom_sf(data = nc_county_sf, aes(fill = 'pink')) +
  geom_sf(data = NorthCarolina_counties_sf, aes(fill = 'white')) + 
  geom_sf(data = Intersection, aes(color = 'red'), size = 0.75) +
  scale_fill_identity(name = "", breaks = c("pink", "white"), 
                      labels = c("North Carolina", "Counties with high hog farm populations"), guide = "legend") +
  scale_color_identity(name = "", breaks = c("red"), 
                       labels = c("Locations of Hog Farms"), guide = "legend") +
  ggtitle("Hog Farms in North Carolina") +
  mytheme
print(HG)
```


# Rationale and Research Questions



\newpage

# Dataset Information



\newpage

# Exploratory Analysis 


=======
## Data Visualization

```{r}

```
>>>>>>> d9cd66887630e0b301f2383a2cc4e03fffe97ba0:JacobChan_FinalProject_HogFarms.Rmd

\newpage

# Analysis
> From our analysis we found......


## Question 1: 

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References

+ [Environmental Working Group Map - Distribution of Animal Agriculture Farms] <https://www.ewg.org/interactive-maps/2020-fields-of-filth/map/>
+ [NC Farms and Floodplains] <https://sites.tufts.edu/gis/files/2017/06/Moore_Emily_UEP232_2016.pdf>
