---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Pigs, Poop, and Public Health"
subtitle: "https://github.com/mchan211/ChanJacob_ENV872_EDA_FinalProject.git"
author: "Natasha Jacob and Meilin Chan"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
#Table of Contents
* List of Tables
* List of Figures 
* Rationale and Research Question
* Dataset Information
* Exploratory Analysis 
  + Data Table
  + Spatial Analysis
  + Visualization Analysis
* Analysis
* Summary and Conclusions
* Resources

\newpage
#List of Tables
* NC County Health Distribution


\newpage
#List of Figures
* [MAP NAME]
* Population Distribution by County (%)
* Chronic Health Incidence Rates
* Health Accessibility 
* Health Risk Indicators

\newpage

# Rationale and Research Questions
> Our main research question was if NC counties with higher concentration of hog farms would have higher health disparities compared to counties with lower hog farm concentrations. NC Hog farms dispose of pig poop by collecting enormous amounts of pig pools into lagoons and spraying the liquidy mixture out onto fields. As you can imagine, the fine mist can travel some distance and has led to strong odors and health impacts of neighboring communities. We were interested in seeing the distribution of populations near where hog farms are concentrated as well as if there were any disparities in health and health access. To research this, we looked at health and population distribution data for each NC county and specifically looked at data for counties with the highest concentrations of hog farms (Bladen, Duplin, Greene, Sampson, Robeson, Wayne) and comparing them to counties with low concentrations or no hog farms (Durham - as we are current residents of this county, Beaufort - as this is part of Duke's community, Martin, Wilson, and Warren). We wanted to gain a better understanding of the inequities and environmental justice issues that stem from NC hog farms and pork production economy. 

\newpage

# Dataset Information
> Our core analysis revolves around the use of spatial datasets and a dataset containing population distribution and health data for each NC county.

* Spatial Data
  + [NAME]
  + [NAME]
* Visualization Data
  + CountyHogs_HealthData.csv


\newpage

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
# Set your working directory
getwd()

# Load your packages
#install.packages("tidyverse")
library(tidyverse)
#install.packages("magrittr")
library(magrittr)
#install.packages("dplyr")
library(dplyr)
#install.packages("sf")
library(sf)
#install.packages("cowplot")
library(cowplot)



# Set your ggplot theme
mytheme <- theme_classic(base_size = 12) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom")
theme_set(mytheme)

# Load your datasets
Health_Data_Raw <- read.csv("../ChanJacob_ENV872_EDA_FinalProject/Data/Raw/County Health Data 2021.csv", stringsAsFactors = TRUE)

```

# Exploratory Analysis 

##Data Table - NC County Health Distribution

```{r, data wrangling, include = FALSE, message = FALSE, error = FALSE}
##Wrangling the dataset 

#Removing columns and rows containing weblinks and other metadata 
Health_Data_Clean <- Health_Data_Raw[-c(1,2,3),-c(1,3, 4, 5, 6)] 

#Using a pipe to select and rename columns
Health_Data_Clean <- Health_Data_Clean %>%
  select(Bold.Indicator.Name.Text, NC, X.6, X.8, X.30,X.31, X.39, X.57, X.81, X.77, X.92, X.95, X.97) %>%
  rename(Indicators = Bold.Indicator.Name.Text, Bladen = X.8, Duplin = X.30, Durham = X.31, Greene = X.39, Sampson = X.81, Robeson = X.77, Wayne = X.95, Wilson = X.97, Warren = X.92, Martin = X.57, Beaufort = X.6)
#Durham - 

#Transposing the data
Health_data_transposed <- t(Health_Data_Clean)

#Converting the transposed data to a dataframe
Health_data_transposed <- as.data.frame(Health_data_transposed)

#Adding a pipe statement to rename the columns
Health_data_transposed_clean <- Health_data_transposed %>%
  rename(Population = 1, Child_Population_percentage = 2, Elderly_Population_percentage = 3, White_Population_percentage = 4, Hispanic_Latinx_percentage = 5, African_American_percentage = 6, American_Indian_percentage = 9, Reading_Proficiency_percentage = 12, Uninsured_Adults_percentage = 14, Medicaid_CHIP_enrolles_percentage = 16, Primary_Care_Physicians = 17, Low_Birthweight_percentage = 32, Infant_Mortality_Rate = 33, Poverty_percentage = 41, Cancer_Incidence = 39, Heart_Disease = 40, Air_Pollution = 52) 

#Selecting the required rows 
Health_data_transposed_clean_1 <- Health_data_transposed_clean[-c(1), -c(7,8, 10, 11, 13, 15, 18:31, 34:38, 42:51, 53:59)]

#Removing '%' character from the dataframe using the 'gsub' function
Health_data_transposed_clean_1 <- Health_data_transposed_clean_1 %>%
  mutate(Population = as.numeric(gsub(",","",Population)),
         Child_Population_percentage = as.numeric(gsub("%", "", Child_Population_percentage)),
         Elderly_Population_percentage = as.numeric(gsub("%", "", Elderly_Population_percentage)),
         White_Population_percentage = as.numeric(gsub("%", "", White_Population_percentage)),
         Hispanic_Latinx_percentage = as.numeric(gsub("%", "", Hispanic_Latinx_percentage)),
         African_American_percentage = as.numeric(gsub("%", "", African_American_percentage)),
         American_Indian_percentage = as.numeric(gsub("%", "", American_Indian_percentage)),
         Reading_Proficiency_percentage = as.numeric(gsub("%", "", Reading_Proficiency_percentage)),
         Uninsured_Adults_percentage = as.numeric(gsub("%", "", Uninsured_Adults_percentage)),
         Medicaid_CHIP_enrolles_percentage = as.numeric(gsub("%", "", Medicaid_CHIP_enrolles_percentage)),
         Low_Birthweight_percentage = as.numeric(gsub("%", "", Low_Birthweight_percentage)),
         Poverty_percentage = as.numeric(gsub("%", "", Poverty_percentage)))



#Checking the class of Health_data_transposed_clean_1
class(Health_data_transposed_clean_1)

#saving wrangled health data in Processed Data folder
write.csv(Health_data_transposed_clean_1, file = "./Data/Processed/projectNCcounties_HealthData.csv")
```

```{r health data frame, echo = FALSE, message = FALSE, error = FALSE}
nccounties.health <- read.csv("./Data/Processed/projectNCcounties_HealthData.csv", stringsAsFactors = FALSE)

print(nccounties.health)
colnames(nccounties.health)
summary(nccounties.health)
```

\newpage

##Spatial Analysis

```{r spatial, error = FALSE, message = FALSE}
mapviewOptions(fgb = FALSE)


#Filtering for NC counties from US counties shapefile
nc_county_sf <- read_sf("./Data/Spatial/cb_2018_us_county_20m.shp") %>%
  filter(STATEFP == 37) #filter for NC counties

#Loading in the shapefile
NorthCarolina_sf <- st_read('../ChanJacob_ENV872_EDA_FinalProject/Data/Spatial/cb_2018_us_county_20m.shp') %>%
  filter(STATEFP == 37)


NorthCarolina_counties_sf <- NorthCarolina_sf %>%
  filter(NAME %in% c("Robeson", "Sampson", "Duplin", "Bladen", "Greene", "Wayne", "Durham"))

mapview(NorthCarolina_counties_sf)

#Loading in the Hog Farm NC locations file 

HogFarm_Locations <- read.csv("../ChanJacob_ENV872_EDA_FinalProject/Data/Spatial/Hog_Farm_Locations_NC.csv") 

#Selecting the required columns with Lat/Long
HogFarm_Locations <- HogFarm_Locations %>%
  select(title, latitude, longitude) %>%
   na.omit(HogFarm_Locations)
  
#Converting the dataframe to a sf (spatial features) dataframe
HogFarm_Locations_sf <- HogFarm_Locations %>%
  st_as_sf(coords = c('longitude', 'latitude'),
           crs = 4269)

colnames(HogFarm_Locations_sf)


HogFarm_Locations_plot <-
ggplot() +
  geom_sf(data = nc_county_sf, aes(fill = 'turquoise')) +
  geom_sf(data = HogFarm_Locations_sf, aes(color = 'red'), size = 1) + 
  scale_fill_identity(name = "", breaks = c("turquoise"), 
                       labels = c("North Carolina"), guide = "legend") +
  scale_color_identity(name = "", breaks = c('red'), 
                       labels = c('Hog Farms'), guide = 'legend') +
  ggtitle("Hog Farms in North Carolina") +
  mytheme 
print(HogFarm_Locations_plot)


Intersection <- HogFarm_Locations_sf[NorthCarolina_counties_sf,]

HG <- 
  ggplot() +
   geom_sf(data = nc_county_sf, aes(fill = 'pink')) +
  geom_sf(data = NorthCarolina_counties_sf, aes(fill = 'white')) + 
  geom_sf(data = Intersection, aes(color = 'red'), size = 0.75) +
  scale_fill_identity(name = "", breaks = c("pink", "white"), 
                      labels = c("North Carolina", "Counties with high hog farm populations"), guide = "legend") +
  scale_color_identity(name = "", breaks = c("red"), 
                       labels = c("Locations of Hog Farms"), guide = "legend") +
  ggtitle("Hog Farms in North Carolina") +
  mytheme
print(HG)
```

\newpage

##Visualization Analysis
###Population
```{r ggplots population, include = FALSE, message = FALSE, error = FALSE}
nccounties.health <- nccounties.health %>% filter(X != "NC")


#POPULATION DISTRIBUTION
NC.pop.distribution <- ggplot(nccounties.health, 
                              aes(reorder(X, Population), 
                                  Population, fill = X)) +
  geom_col() +
  xlab("") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
NC.pop.distribution

w.pop.distribution <- ggplot(nccounties.health, 
                             aes(reorder(X, White_Population_percentage), 
                                 White_Population_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("White") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

aa.pop.distribution <- ggplot(nccounties.health, 
                              aes(reorder(X, African_American_percentage),
                                  African_American_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("African American")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

h.pop.distribution <- ggplot(nccounties.health,
                             aes(reorder(X, Hispanic_Latinx_percentage),
                                 Hispanic_Latinx_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Hispanic/Latinx")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

i.pop.distribution <- ggplot(nccounties.health, 
                             aes(reorder(X, American_Indian_percentage),
                                 American_Indian_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Indigenous/Native American")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

counties.legend <- get_legend(NC.pop.distribution)

```

```{r pop final, echo = FALSE, message = FALSE, error = FALSE}
pop.plot <- plot_grid(NC.pop.distribution + theme(legend.position = "none"),
                      w.pop.distribution + theme(legend.position = "none"),
                      aa.pop.distribution + theme(legend.position = "none"),
                      h.pop.distribution + theme(legend.position = "none"),
                      i.pop.distribution + theme(legend.position = "none"),
                      nrow = 2,
                      align = 'hv')
pop.title <- ggdraw() +
  draw_label("Population Distribution by County(%)",
             fontface='bold',
             x = 0,
             hjust = 0) +
  theme(plot.margin = margin (0,0,0,7))

pop.plot.complete <- plot_grid(pop.title, pop.plot, 
                               ncol = 1, rel_heights = c(0.1, 1))

pop.plot.complete

```
\newpage

###Chronic Illness
```{r ggplots disease, include = FALSE, message = FALSE, error = FALSE}
#Chronic Health
cancer.plot <- ggplot(nccounties.health, 
                      aes(reorder(X, Cancer_Incidence),
                          Cancer_Incidence, fill = X)) +
  geom_col() +
  xlab("")+
  ylab("Cancer Incidence Rates per 100,000 pop.")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

heart.plot <- ggplot(nccounties.health,
                     aes(reorder(X, Heart_Disease),
                         Heart_Disease, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Heart Disease Deaths per 100,000 pop.")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r disease final, echo = FALSE, message = FALSE, error = FALSE}
chronic.disease.plot <- plot_grid(cancer.plot + theme(legend.position = "none"),
                                  heart.plot + theme(legend.position = "none"),
                                  nrow = 1,
                                  align = 'hv')
disease.title <- ggdraw() +
  draw_label("Chronic Health Incidence Rates",
             fontface='bold',
             x = 0,
             hjust = 0) +
  theme(plot.margin = margin (0,0,0,7))

chronic.disease.complete <- plot_grid(disease.title, chronic.disease.plot,
                                      ncol = 1, rel_heights = c(0.1, 1))
chronic.disease.complete
```
\newpage
###Health Challenge Indicators
```{r, health plot, include = FALSE, message = FALSE, error = FALSE}
phys.plot <- ggplot(nccounties.health,
                    aes(reorder(X, Primary_Care_Physicians),
                        Primary_Care_Physicians, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Physicians per 10,000 pop.")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

pov.plot <- ggplot(nccounties.health,
                   aes(reorder(X, Poverty_percentage),
                       Poverty_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% Individ. Living in Poverty") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
  
air.plot <- ggplot(nccounties.health,
                   aes(reorder(X, Air_Pollution),
                       Air_Pollution, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Avg. Daily Density of PM (u/m3)") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

uninsured.plot <- ggplot(nccounties.health,
                         aes(reorder(X, Uninsured_Adults_percentage),
                             Uninsured_Adults_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% Uninsured Adults") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

medicaid.plot <- ggplot(nccounties.health, 
                        aes(reorder(X, Medicaid_CHIP_enrolles_percentage),
                            Medicaid_CHIP_enrolles_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% Pop. Enrolled in Medicaid") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

```{r, final health, echo = FALSE, message = FALSE, error = FALSE}
health.plot.1 <- plot_grid (phys.plot + theme(legend.position = "none",
                                              axis.title = element_text(size = 7)),
                          pov.plot + theme(legend.position = "none",
                                           axis.title = element_text(size = 7)),
                          air.plot + theme(legend.position = "none",
                                           axis.title = element_text(size = 7)),
                          ncol = 3,
                          align = 'hv')
health.plot.2 <- plot_grid( uninsured.plot + theme(legend.position = "none",
                                                   axis.title = element_text(size = 7)),
                          medicaid.plot + theme(legend.position = "none",
                                                axis.title = element_text(size = 7)),
                          ncol = 2,
                          align = 'hv')

health.plot <- plot_grid(health.plot.1, health.plot.2,
                         nrow = 2,
                         align = 'hv')
health.title <- ggdraw() +
  draw_label("Health Risk Indicators",
             fontface='bold',
             x = 0,
             hjust = 0) +
  theme(plot.margin = margin (0,0,0,7))

health.plot.final <- plot_grid(health.title, health.plot,
                               nrow = 2, rel_heights = c(0.1, 1))
health.plot.final

```
\newpage
###Infant Health
```{r infant plots, include = FALSE, message = FALSE, error = FALSE}
birthweight.plot <- ggplot(nccounties.health,
                           aes(reorder(X, Low_Birthweight_percentage),
                               Low_Birthweight_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% newborns under 2.5kg")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

mortality.plot <- ggplot(nccounties.health,
                         aes(reorder(X, Infant_Mortality_Rate),
                             Infant_Mortality_Rate, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Infant Death Rates per 1,000 births")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r final infant, echo = FALSE, message = FALSE, error = FALSE}
infant.plot <- plot_grid(birthweight.plot + theme(legend.position = "none"),
                         mortality.plot + theme(legend.position = "none"),
                         ncol = 2,
                         align = 'hv')

infant.title <- ggdraw() +
  draw_label("Infant Health Disparities",
             fontface='bold',
             x = 0,
             hjust = 0) +
  theme(plot.margin = margin (0,0,0,7))

infant.plot.final <- plot_grid(infant.title, infant.plot,
                               ncol = 1, rel_heights = c(0.1, 1))
infant.plot.final

```

\newpage
###Vulnerable Populations
```{r vulnerable, include = FALSE, message = FALSE, error = FALSE}

chil.plot <- ggplot(nccounties.health,
                           aes(reorder(X, Child_Population_percentage),
                               Child_Population_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% Population < 18 Years")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

el.plot <- ggplot(nccounties.health,
                           aes(reorder(X, Elderly_Population_percentage),
                               Elderly_Population_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% Population 65+ Years")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

```{r vulnerable final, echo = FALSE, mesage = FALSE, error = FALSE}

vul.plot <- plot_grid(chil.plot + theme(legend.position = "none"),
                      el.plot + theme(legend.position = "none"),
                      ncol = 2,
                      align = 'hv')
vul.title <- ggdraw() +
  draw_label("Vulnerable Populations",
             fontface='bold',
             x = 0,
             hjust = 0) +
  theme(plot.margin = margin (0,0,0,7))

vul.plot.final <- plot_grid(vul.title, vul.plot,
                            ncol = 1, rel_heights = c(0.1, 1))
vul.plot.final

```



\newpage

#Analysis


## Question 1: 

## Question 2: 



\newpage

# Summary and Conclusions


\newpage

# References

* [Environmental Working Group Map - Distribution of Animal Agriculture Farms] <https://www.ewg.org/interactive-maps/2020-fields-of-filth/map/>
* [NC Farms and Floodplains] <https://sites.tufts.edu/gis/files/2017/06/Moore_Emily_UEP232_2016.pdf>
* [Democracy Now! Video on NC Hog Farms] <https://www.youtube.com/watch?v=eyAFNV4Afgw>
