---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Insert title of project here"
subtitle: "Web address for GitHub repository"
author: "Name"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
getwd()

# Load your packages
#install.packages("tidyverse")
library(tidyverse)
#install.packages("magrittr")
library(magrittr)
#install.packages("dplyr")
library(dplyr)
library(sf)

# Set your ggplot theme
mytheme <- theme_classic(base_size = 12) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

# Load your datasets
Health_Data_Raw <- read.csv("../ChanJacob_ENV872_EDA_FinalProject/Data/Raw/County Health Data 2021.csv", stringsAsFactors = TRUE)

```

#Analysis

##Data Wrangling

```{r}
##Wrangling the dataset 

#Removing columns and rows containing weblinks and other metadata 
Health_Data_Clean <- Health_Data_Raw[-c(1,2,3),-c(1,3, 4, 5, 6)] 

#Using a pipe to select and rename columns
Health_Data_Clean <- Health_Data_Clean %>%
  select(Bold.Indicator.Name.Text, NC, X.8, X.30,X.31, X.39, X.81, X.77, X.95) %>%
  rename(Indicators = Bold.Indicator.Name.Text, Bladen = X.8, Duplin = X.30, Durham = X.31, Greene = X.39, Sampson = X.81, Robeson = X.77, Wayne = X.95)
#Durham - 

#Transposing the data
Health_data_transposed <- t(Health_Data_Clean)

#Converting the transposed data to a dataframe
Health_data_transposed <- as.data.frame(Health_data_transposed)

#Adding a pipe statement to rename the columns
Health_data_transposed_clean <- Health_data_transposed %>%
  rename(Population = 1, Child_Population_percentage = 2, Elderly_Population_percentage = 3, White_Population_percentage = 4, Hispanic_Latinx_percentage = 5, African_American_percentage = 6, American_Indian_percentage = 9, Reading_Proficiency_percentage = 12, Uninsured_Adults_percentage = 14, Medicaid_CHIP_enrolles_percentage = 16, Primary_Care_Physicians = 17, Low_Birthweight_percentage = 32, Infant_Mortality_Rate = 33, Poverty_percentage = 41, Air_Pollution = 52) 

#Selecting the required rows 
Health_data_transposed_clean_1 <- Health_data_transposed_clean[-c(1), -c(7,8, 10, 11, 13, 15, 18:31, 34:40, 42:51, 53:59)]

#Removing '%' character from the dataframe using the 'gsub' function
Health_data_transposed_clean_1 <- Health_data_transposed_clean_1 %>%
  mutate(Child_Population_percentage = as.numeric(gsub("%", "", Child_Population_percentage)),
         Elderly_Population_percentage = as.numeric(gsub("%", "", Elderly_Population_percentage)),
         White_Population_percentage = as.numeric(gsub("%", "", White_Population_percentage)),
         Hispanic_Latinx_percentage = as.numeric(gsub("%", "", Hispanic_Latinx_percentage)),
         African_American_percentage = as.numeric(gsub("%", "", African_American_percentage)),
         American_Indian_percentage = as.numeric(gsub("%", "", American_Indian_percentage)),
         Reading_Proficiency_percentage = as.numeric(gsub("%", "", Reading_Proficiency_percentage)),
         Uninsured_Adults_percentage = as.numeric(gsub("%", "", Uninsured_Adults_percentage)),
         Medicaid_CHIP_enrolles_percentage = as.numeric(gsub("%", "", Medicaid_CHIP_enrolles_percentage)),
         Low_Birthweight_percentage = as.numeric(gsub("%", "", Low_Birthweight_percentage)),
         Poverty_percentage = as.numeric(gsub("%", "", Poverty_percentage)))

#Filtering for NC counties from US counties shapefile
nc_county_sf <- read_sf("./Data/Spatial/cb_2018_us_county_20m.shp") %>%
  filter(STATEFP == 37) #filter for NC counties


```

## Data Visualization

```{r}

```





# Rationale and Research Questions
> Our main research question was if NC counties with higher concentration of hog farms would have higher health disparities compared to counties with lower hog farm concentrations. To research this, we looked at health and population distribution data for each NC county and specifically looked at data for counties with the highest concentrations of hog farms (Bladen, Duplin, Greene, Sampson, Robeson, Wayne) and comparing them to Durham as we are current residents of this county. We wanted to gain a better understanding of the inequities and environmental justice issues that stem from NC hog farms and pork production economy. 


\newpage

# Dataset Information



\newpage

# Exploratory Analysis 



\newpage

# Analysis



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
+ [Environmental Working Group Map - Distribution of Animal Agriculture Farms] <https://www.ewg.org/interactive-maps/2020-fields-of-filth/map/>
+ [NC Farms and Floodplains] <https://sites.tufts.edu/gis/files/2017/06/Moore_Emily_UEP232_2016.pdf>
