---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Pigs, Poop, and Public Health"
subtitle: "https://github.com/mchan211/ChanJacob_ENV872_EDA_FinalProject.git"
author: "Natasha Jacob and Meilin Chan"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
# Table of Contents
- List of Tables
- List of Figures 
- Rationale and Research Question
- Dataset Information
- Analysis 
  - Data Table
  - Spatial Analysis
  - Visualization Analysis
- Summary and Conclusions
- Resources

\newpage
# List of Tables
- NC County Health Distribution
- Counties with high populations of hog farms

\newpage
# List of Figures
- Hog Farm locations in North Carolina
- Counties with High Populations of Hog Farms
- Population Distribution by County (%)
- Chronic Health Incidence Rates
- Health Accessibility 
- Health Risk Indicators

\newpage

# Introduction
> Hog farming is the practice of raising and breeding domestic pigs mainly for their meat and skin. Pigs are a popular form of livestock, with more than one billion pigs butchered each year worldwide, 100 million of them in the USA. The discharge of waste from hog farms is usually done by flushing the waste into giant holding ponds called lagoons. The liquid waste from the lagoons is then sprayed onto the farmers’ fields to keep the lagoons from filling. The result, according to the farms’ neighbors, is stench and insects and decay from dead hogs that are piled into open containers before they are removed. The EPA and federal politicians have drafted little to no regulation surrounding CAFOs to safeguard the environment and human welfare from their effects. Because manure lagoons are not connected to a moving water supply, there are little to no regulations surrounding waste disposal. As a result, they are not considered a serious threat to human or environmental health. 

# Rationale and Research Questions
> Our main research question was if NC counties with higher concentration of hog farms would have higher health disparities compared to counties with lower hog farm concentrations. NC Hog farms dispose of pig poop by collecting enormous amounts of pig pools into lagoons and spraying the liquidy mixture out onto fields. As you can imagine, the fine mist can travel some distance and has led to strong odors and health impacts of neighboring communities. We were interested in seeing the distribution of populations near where hog farms are concentrated as well as if there were any disparities in health and health access. To research this, we looked at health and population distribution data for each NC county and specifically looked at data for counties with the highest concentrations of hog farms (Bladen, Duplin, Greene, Sampson, Robeson, Wayne) and comparing them to counties with low concentrations or no hog farms (Durham - as we are current residents of this county, Beaufort - as this is part of Duke's community, Martin, Wilson, and Warren). We wanted to gain a better understanding of the inequities and environmental justice issues that stem from NC hog farms and pork production economy. 

\newpage

# Dataset Information
> Our core analysis revolves around the use of spatial datasets of NC counties and hog farm locations, and a dataset containing population distribution and health data for each NC county.

* Spatial Data
  + cb_2018_us_county_20m.shp
  + Hog_Farm_Locations_NC1.csv
* Visualization Data
  + CountyHogs_HealthData.csv


\newpage

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
# Set your working directory
getwd()

# Load your packages
#install.packages("tidyverse")
library(tidyverse)
#install.packages("magrittr")
library(magrittr)
#install.packages("dplyr")
library(dplyr)
#install.packages("sf")
library(sf)
#install.packages("cowplot")
library(cowplot)

# Load your datasets
Health_Data_Raw <- read.csv("../ChanJacob_ENV872_EDA_FinalProject/Data/Raw/County Health Data 2021.csv", stringsAsFactors = TRUE)

```

# Analysis 

##Data Table - NC County Health Distribution

```{r error=FALSE, message=FALSE, include=FALSE}
##Wrangling the dataset 
#Removing columns and rows containing weblinks and other metadata 
Health_Data_Clean <- Health_Data_Raw[-c(1,2,3),-c(1,3, 4, 5, 6)] 

#Using a pipe to select and rename columns
Health_Data_Clean <- Health_Data_Clean %>%
  select(Bold.Indicator.Name.Text, NC, X.6, X.8, X.30,X.31, X.39, X.57, X.81, X.77, X.92, X.95, X.97) %>%
  rename(Indicators = Bold.Indicator.Name.Text, Bladen = X.8, Duplin = X.30, Durham = X.31, Greene = X.39, Sampson = X.81, Robeson = X.77, Wayne = X.95, Wilson = X.97, Warren = X.92, Martin = X.57, Beaufort = X.6)

#Transposing the data
Health_data_transposed <- t(Health_Data_Clean)

#Converting the transposed data to a dataframe
Health_data_transposed <- as.data.frame(Health_data_transposed)

#Adding a pipe statement to rename the columns
Health_data_transposed_clean <- Health_data_transposed %>%
  rename(Population = 1, Child_Population_percentage = 2, Elderly_Population_percentage = 3, White_Population_percentage = 4, Hispanic_Latinx_percentage = 5, African_American_percentage = 6, American_Indian_percentage = 9, Reading_Proficiency_percentage = 12, Uninsured_Adults_percentage = 14, Medicaid_CHIP_enrolles_percentage = 16, Primary_Care_Physicians = 17, Low_Birthweight_percentage = 32, Infant_Mortality_Rate = 33, Poverty_percentage = 41, Cancer_Incidence = 39, Heart_Disease = 40, Air_Pollution = 52) 

#Selecting the required rows 
Health_data_transposed_clean_1 <- Health_data_transposed_clean[-c(1), -c(7,8, 10, 11, 13, 15, 18:31, 34:38, 42:51, 53:59)]

#Removing '%' character from the dataframe using the 'gsub' function
Health_data_transposed_clean_1 <- Health_data_transposed_clean_1 %>%
  mutate(Population = as.numeric(gsub(",","",Population)),
         Child_Population_percentage = as.numeric(gsub("%", "", Child_Population_percentage)),
         Elderly_Population_percentage = as.numeric(gsub("%", "", Elderly_Population_percentage)),
         White_Population_percentage = as.numeric(gsub("%", "", White_Population_percentage)),
         Hispanic_Latinx_percentage = as.numeric(gsub("%", "", Hispanic_Latinx_percentage)),
         African_American_percentage = as.numeric(gsub("%", "", African_American_percentage)),
         American_Indian_percentage = as.numeric(gsub("%", "", American_Indian_percentage)),
         Reading_Proficiency_percentage = as.numeric(gsub("%", "", Reading_Proficiency_percentage)),
         Uninsured_Adults_percentage = as.numeric(gsub("%", "", Uninsured_Adults_percentage)),
         Medicaid_CHIP_enrolles_percentage = as.numeric(gsub("%", "", Medicaid_CHIP_enrolles_percentage)),
         Low_Birthweight_percentage = as.numeric(gsub("%", "", Low_Birthweight_percentage)),
         Poverty_percentage = as.numeric(gsub("%", "", Poverty_percentage)))



#Checking the class of Health_data_transposed_clean_1
class(Health_data_transposed_clean_1)

#saving wrangled health data in Processed Data folder
write.csv(Health_data_transposed_clean_1, file = "./Data/Processed/projectNCcounties_HealthData.csv")
```


```{r health data frame, error=FALSE, message=FALSE, include=FALSE}
nccounties.health <- read.csv("./Data/Processed/projectNCcounties_HealthData.csv", stringsAsFactors = FALSE)

print(nccounties.health)
colnames(nccounties.health)
summary(nccounties.health)
```

\newpage

## Spatial Analysis

<<<<<<< HEAD
```{r spatial, error = FALSE, message = FALSE}
#mapviewOptions(fgb = FALSE)
=======
```{r message=FALSE, warning=FALSE, include=FALSE}
#Loading in the shapefile and filtering for NC counties
NorthCarolina_sf <- st_read('../ChanJacob_ENV872_EDA_FinalProject/Data/Spatial/cb_2018_us_county_20m.shp') %>%
  filter(STATEFP == 37)

```

**Hog Farms in North Carolina**
```{r spatial, echo=FALSE, error=FALSE, fig.height=7, fig.width=8, message=FALSE, fig.align='left'}
>>>>>>> 2b12c73176e1ff209d72c90dddf3b262d6cfa4b0

# Set your ggplot theme
mytheme <- theme_classic(base_size = 12) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        line = element_blank())
theme_set(mytheme)

#Filtering counties that have a high population of Hog farms
HighPop_HogFarms_sf <- NorthCarolina_sf %>%
  filter(NAME %in% c("Robeson", "Sampson", "Duplin", "Bladen", "Greene", "Wayne"))

#Filtering counties that have no or low populations of Hog farms
LowPop_HogFarms_sf <- NorthCarolina_sf %>%
  filter(NAME %in% c("Wilson", "Martin", "Beaufort", "Durham", "Warren"))


#Loading in the Hog Farm NC locations file 
HogFarm_Locations <- read.csv("../ChanJacob_ENV872_EDA_FinalProject/Data/Spatial/Hog_Farm_Locations_NC1.csv") 

#Selecting the required columns with Lat/Long
HogFarm_Locations <- HogFarm_Locations %>%
  select(title, latitude, longitude) %>%
   na.omit(HogFarm_Locations)
  
#Converting the dataframe to a sf (spatial features) dataframe
HogFarm_Locations_sf <- HogFarm_Locations %>%
  st_as_sf(coords = c('longitude', 'latitude'),
           crs = 4269)

#Spatial Analysis of Hog Farms in all of NC
HogFarm_Locations_plot <-
ggplot() +
  geom_sf(data = NorthCarolina_sf, aes(fill = 'white')) +
  geom_sf(data = HighPop_HogFarms_sf, aes(fill = 'lightpink')) + 
  geom_sf(data = LowPop_HogFarms_sf, aes(fill = 'powderblue')) +
  geom_sf(data = HogFarm_Locations_sf, aes(color = 'blue'), size = 2, shape = 8) +
  scale_fill_identity(name = "", breaks = c("white", "lightpink", "powderblue"), 
                       labels = c("North Carolina", "High density Hog Farms", "Low density Hog Farms"), guide = "legend") +
  scale_color_identity(name = "", breaks = c('blue'), 
                       labels = c('Hog Farm Locations'), guide = 'legend') +
  #ggtitle("Hog Farms in North Carolina") +
  mytheme 
print(HogFarm_Locations_plot)



```

**NC counties with high populations of Hog Farms**

```{r echo=FALSE, error=FALSE, message=FALSE, fig.align='left'}

#Spatial Analysis of Counties with a High Population of Hog Farms 
HighPop_HogFarms_sf <- NorthCarolina_sf %>%
  filter(NAME %in% c("Robeson", "Sampson", "Duplin", "Bladen", "Greene", "Wayne"))


Robeson_sf <- NorthCarolina_sf %>%
  filter(NAME %in% c("Robeson"))


Sampson_sf <- NorthCarolina_sf %>%
  filter(NAME %in% c("Sampson"))

Duplin_sf <- NorthCarolina_sf %>%
  filter(NAME %in% c("Duplin"))


Bladen_sf <- NorthCarolina_sf %>%
  filter(NAME %in% c("Bladen"))

Greene_sf <- NorthCarolina_sf %>%
  filter(NAME %in% c("Greene"))


Wayne_sf <- NorthCarolina_sf %>%
  filter(NAME %in% c("Wayne"))

#Finding out the no. of hog farms in each county
Inter_Rob <- HogFarm_Locations_sf[Robeson_sf,]
Inter_Sam <- HogFarm_Locations_sf[Sampson_sf,]
Inter_Dup <- HogFarm_Locations_sf[Duplin_sf,]
Inter_Blad <- HogFarm_Locations_sf[Bladen_sf,]
Inter_Greene <- HogFarm_Locations_sf[Greene_sf,]


Intersection <- HogFarm_Locations_sf[HighPop_HogFarms_sf,] 
 
HighPop <-
ggplot() +
  geom_sf(data = Robeson_sf, aes(fill = 'lavender')) + 
  geom_sf(data = Sampson_sf, aes(fill = 'lightblue1')) +
  geom_sf(data = Duplin_sf, aes(fill = 'palegreen')) +
  geom_sf(data = Bladen_sf, aes(fill = 'thistle1')) +
  geom_sf(data = Greene_sf, aes(fill = 'lightyellow1')) +
  geom_sf(data = Wayne_sf, aes(fill = 'rosybrown1')) +
  geom_sf(data = Intersection, aes(color = 'blue'), size = 2, shape = 8) +
  scale_fill_identity(name = "", breaks = c("lavender", "lightblue1", "palegreen", "thistle1", "lightyellow1", "rosybrown1"),
                      labels = c('Robeson', 'Sampson', 'Duplin', 'Bladen', 'Greene', 'Wayne'),
                      guide = "legend") +
    scale_color_identity(name = "", breaks = c("blue"), labels = c("Hog Farm Locations"), guide = "legend") +
  #ggtitle("NC counties with high populations of Hog Farms") +
  mytheme
print(HighPop)
```


\newpage

## Visualization Analysis
## Population
```{r ggplots population, include = FALSE, message = FALSE, error = FALSE}
nccounties.health <- nccounties.health %>% filter(X != "NC")


#POPULATION DISTRIBUTION
NC.pop.distribution <- ggplot(nccounties.health, 
                              aes(reorder(X, Population), 
                                  Population, fill = X)) +
  geom_col() +
  xlab("") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
NC.pop.distribution

w.pop.distribution <- ggplot(nccounties.health, 
                             aes(reorder(X, White_Population_percentage), 
                                 White_Population_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("White") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

aa.pop.distribution <- ggplot(nccounties.health, 
                              aes(reorder(X, African_American_percentage),
                                  African_American_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("African American")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

h.pop.distribution <- ggplot(nccounties.health,
                             aes(reorder(X, Hispanic_Latinx_percentage),
                                 Hispanic_Latinx_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Hispanic/Latinx")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

i.pop.distribution <- ggplot(nccounties.health, 
                             aes(reorder(X, American_Indian_percentage),
                                 American_Indian_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Indigenous/Native American")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

counties.legend <- get_legend(NC.pop.distribution)

```

```{r pop final, echo = FALSE, message = FALSE, error = FALSE}
pop.plot <- plot_grid(NC.pop.distribution + theme(legend.position = "none"),
                      w.pop.distribution + theme(legend.position = "none"),
                      aa.pop.distribution + theme(legend.position = "none"),
                      h.pop.distribution + theme(legend.position = "none"),
                      i.pop.distribution + theme(legend.position = "none"),
                      nrow = 2,
                      align = 'hv')
pop.title <- ggdraw() +
  draw_label("Population Distribution by County(%)",
             fontface='bold',
             x = 0,
             hjust = 0) +
  theme(plot.margin = margin (0,0,0,7))

pop.plot.complete <- plot_grid(pop.title, pop.plot, 
                               ncol = 1, rel_heights = c(0.1, 1))

pop.plot.complete

```
\newpage

###Chronic Illness
```{r ggplots disease, include = FALSE, message = FALSE, error = FALSE}
#Chronic Health
cancer.plot <- ggplot(nccounties.health, 
                      aes(reorder(X, Cancer_Incidence),
                          Cancer_Incidence, fill = X)) +
  geom_col() +
  xlab("")+
  ylab("Cancer Incidence Rates per 100,000 pop.")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

heart.plot <- ggplot(nccounties.health,
                     aes(reorder(X, Heart_Disease),
                         Heart_Disease, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Heart Disease Deaths per 100,000 pop.")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r disease final, echo = FALSE, message = FALSE, error = FALSE}
chronic.disease.plot <- plot_grid(cancer.plot + theme(legend.position = "none"),
                                  heart.plot + theme(legend.position = "none"),
                                  nrow = 1,
                                  align = 'hv')
disease.title <- ggdraw() +
  draw_label("Chronic Health Incidence Rates",
             fontface='bold',
             x = 0,
             hjust = 0) +
  theme(plot.margin = margin (0,0,0,7))

chronic.disease.complete <- plot_grid(disease.title, chronic.disease.plot,
                                      ncol = 1, rel_heights = c(0.1, 1))
chronic.disease.complete
```
\newpage
###Health Challenge Indicators
```{r, health plot, include = FALSE, message = FALSE, error = FALSE}
phys.plot <- ggplot(nccounties.health,
                    aes(reorder(X, Primary_Care_Physicians),
                        Primary_Care_Physicians, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Physicians per 10,000 pop.")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

pov.plot <- ggplot(nccounties.health,
                   aes(reorder(X, Poverty_percentage),
                       Poverty_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% Individ. Living in Poverty") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
  
air.plot <- ggplot(nccounties.health,
                   aes(reorder(X, Air_Pollution),
                       Air_Pollution, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Avg. Daily Density of PM (u/m3)") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

uninsured.plot <- ggplot(nccounties.health,
                         aes(reorder(X, Uninsured_Adults_percentage),
                             Uninsured_Adults_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% Uninsured Adults") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

medicaid.plot <- ggplot(nccounties.health, 
                        aes(reorder(X, Medicaid_CHIP_enrolles_percentage),
                            Medicaid_CHIP_enrolles_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% Pop. Enrolled in Medicaid") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

```{r, final health, echo = FALSE, message = FALSE, error = FALSE}
health.plot.1 <- plot_grid (phys.plot + theme(legend.position = "none",
                                              axis.title = element_text(size = 7)),
                          pov.plot + theme(legend.position = "none",
                                           axis.title = element_text(size = 7)),
                          air.plot + theme(legend.position = "none",
                                           axis.title = element_text(size = 7)),
                          ncol = 3,
                          align = 'hv')
health.plot.2 <- plot_grid( uninsured.plot + theme(legend.position = "none",
                                                   axis.title = element_text(size = 7)),
                          medicaid.plot + theme(legend.position = "none",
                                                axis.title = element_text(size = 7)),
                          ncol = 2,
                          align = 'hv')

health.plot <- plot_grid(health.plot.1, health.plot.2,
                         nrow = 2,
                         align = 'hv')
health.title <- ggdraw() +
  draw_label("Health Risk Indicators",
             fontface='bold',
             x = 0,
             hjust = 0) +
  theme(plot.margin = margin (0,0,0,7))

health.plot.final <- plot_grid(health.title, health.plot,
                               nrow = 2, rel_heights = c(0.1, 1))
health.plot.final

```

\newpage
###Infant Health
```{r infant plots, include = FALSE, message = FALSE, error = FALSE}
birthweight.plot <- ggplot(nccounties.health,
                           aes(reorder(X, Low_Birthweight_percentage),
                               Low_Birthweight_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% newborns under 2.5kg")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

mortality.plot <- ggplot(nccounties.health,
                         aes(reorder(X, Infant_Mortality_Rate),
                             Infant_Mortality_Rate, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("Infant Death Rates per 1,000 births")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r final infant, echo = FALSE, message = FALSE, error = FALSE}
infant.plot <- plot_grid(birthweight.plot + theme(legend.position = "none"),
                         mortality.plot + theme(legend.position = "none"),
                         ncol = 2,
                         align = 'hv')

infant.title <- ggdraw() +
  draw_label("Infant Health Disparities",
             fontface='bold',
             x = 0,
             hjust = 0) +
  theme(plot.margin = margin (0,0,0,7))

infant.plot.final <- plot_grid(infant.title, infant.plot,
                               ncol = 1, rel_heights = c(0.1, 1))
infant.plot.final

```

\newpage
###Vulnerable Populations
```{r vulnerable, include = FALSE, message = FALSE, error = FALSE}

chil.plot <- ggplot(nccounties.health,
                           aes(reorder(X, Child_Population_percentage),
                               Child_Population_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% Population < 18 Years")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

el.plot <- ggplot(nccounties.health,
                           aes(reorder(X, Elderly_Population_percentage),
                               Elderly_Population_percentage, fill = X)) +
  geom_col() +
  xlab("") +
  ylab("% Population 65+ Years")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

```{r vulnerable final, echo = FALSE, mesage = FALSE, error = FALSE}

vul.plot <- plot_grid(chil.plot + theme(legend.position = "none"),
                      el.plot + theme(legend.position = "none"),
                      ncol = 2,
                      align = 'hv')
vul.title <- ggdraw() +
  draw_label("Vulnerable Populations",
             fontface='bold',
             x = 0,
             hjust = 0) +
  theme(plot.margin = margin (0,0,0,7))

vul.plot.final <- plot_grid(vul.title, vul.plot,
                            ncol = 1, rel_heights = c(0.1, 1))
vul.plot.final

```



\newpage

#Analysis

## Question 1: Which NC counties have the highest concentration of hog farms?
> Counties with the highest concentration of hog farms in North Carolina are Bladen, Duplin, Greene, Robeson, Sampson and Wayne. 

|    County      |    Number of Hog Farms     |
|:--------------:|:--------------------------:|
|  Sampson       |          28                |
|  Duplin        |          25                |
|  Bladen        |          19                |
|  Robeson       |          6                 |
|  Greene        |          4                 |

## Question 2: Are hog farms more prevelant in communities with higher BIPOC populations?
> Counties with higher concentrations of hog farms tend to have a higher distribution of Hispanic/Latinx and Native American communities compared to counties with low concentrations of or no hog farms. In particular, Robeson county has the highest proportion of Native American individuals compared to all the other counties we analyzed. Sampson and Duplin have the highest number of hog farms compared to all the counties we identified, and also has the highest proportion of Hispanic/Latinx individuals compared to the counties we analyzed. 

## Question 3: What health inequities are most prevelant in counties with  high concentratiosn of hog farms?
> Counties with high concentrations of hog farms tended to have lower number of physicians per 10,000 people compared to the other counties we observed. Further, these counties also had higher % of adults without health insurance. Robeson county, which also had the highest distribution of Native American individuals, also had the highest % of its population living in poverty. Counties with higher concentrations of hog farms also had slightly higher air PM density measurements - it should be noted that Durham is considered metropolitan compared to the other analyzed counties and thus would have a higher average density of PM. The counties with the highest concentrations of hog farms in particular (Sampson, Duplin) has a population where a quarter is under 18 years old which is a population vulnerable to environmental health impacts. 


\newpage

# Summary and Conclusions
> Our analysis was just a surface level look at potential connections between hog farm location in NC, BIPOC communities, and health disparities. From our visualizations, we see that higher concentrations of hog farms tend to be in counties with higher distributions of Hispanic/Latinx and Native American communities compared to counties with low or no hog farms. Further, communities with higher concentrations of hog farms also tended to have lower numbers of available physicians per population size, higher % of adults without health insurance, higher average daily density PM measurements, and higher % population of children (under 18 years old). Although our analysis did not find strong relationships between counties with high concentrations of hog farms and public health impacts on chronic diseases and infant health, it is possible there are specific health impacts that are unreported. Our data set did not provide data on respiratory health or other less deathly health impacts (ie: headaches, nausea, etc.) that individuals living near hog farms could be experiencing more frequently. 

> Future studies could dig further into the impact of hog farm feces disposal methods on air quality and water quality of Wayne, Bladen, Greene, Robeson, Sampson, and Duplin counties. With air quality, a study could measure and visualize daily air quality measurements and see if there are any increases in PM specifically during periods where the lagoon feces mixture is sprayed into the air. One could also look into if there are any unreported health impacts due to inaccessability to health care in counties with high concentrations of hog farms (ie: no health insurance, low number of physicians compared to population). 

\newpage

# References

* [Environmental Working Group Map - Distribution of Animal Agriculture Farms] <https://www.ewg.org/interactive-maps/2020-fields-of-filth/map/>
* [NC Farms and Floodplains] <https://sites.tufts.edu/gis/files/2017/06/Moore_Emily_UEP232_2016.pdf>
* [Democracy Now! Video on NC Hog Farms] <https://www.youtube.com/watch?v=eyAFNV4Afgw>
* [Cartographic Boundary File] <https://www.arcgis.com/sharing/rest/content/items/05f6d4797e2a428d96c15aba40088159/info/metadata/metadata.xml?format=default&output=html>
* [Locations of Hog Farms in North Carolina] <https://www.google.com/maps/search/hog+farms+in+north+carolina/@35.2918991,-79.6807785,8z/data=!3m1!4b1>
* [NC Health News] <https://www.northcarolinahealthnews.org/2020/12/04/neighbors-of-hog-farms-say-recent-appeals-court-ruling-gives-them-hope/>
* [Phantom Buster] <https://phantombuster.com/>
* [Environmental Health and impacts of pig farming] <https://en.wikipedia.org/wiki/Pig_farming#Environmental_and_health_impacts>
