---
title: "ChanJacob_README"
author: "Meilin Chan and Natasha Jacob"
date: "4/6/2022"
output: pdf_document
---

# <ChanJacob_ENV872_EDA_FinalProject>
```{r setup, include = FALSE, message = FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(sf)
library(mapview)
library(RColorBrewer)



```

<Instructions: copy and paste this template into your project README file (found in the parent folder of the repository). Fill in relevant information as requested.>

<General notes: add as much information as is relevant for your repository. Some overarching guidelines are provided, but feel free to expand on these guidelines.>
<More resources found here: https://www.dataone.org/all-best-practices>
<Delete the text inside the brackets when formatting your file.>

## Summary

> The purpose of our repository is to gather relevant data and analyze the potential connection between locations of North Carolina hog farms, public health issues, and community demographics. 
We will compare racial, health, etc. distributions within North Carolina compared to counties with the highest concentration of hog farms. 

>(4) population, (5) child pop, (6) elderly pop,(7) White, (8)Hispanic/Latinx, (9)African American, (12)American Indian, (15)Reading Proficiency, (17)Uninsured Adults, (19) Medicaid & CHIP Enrolees, (20)Ratio primary care, (35) Low Birthweight, (36)Infant Mortality Rate, (44) Poverty, (55) Air Pollution

## Investigators

**Meilin Chan**  
*meilin.chan@duke.edu*  
  
**Natasha Jacob**  
*natasha.jacob@duke.edu*  


## Keywords

* hog farms
* environmental racism
* environmental health
* health disparities

## Database Information

* County Health Data 2021.xlsx (Raw folder)
  + Source: [NCIOM Data] <https://nciom.org/nc-health-data/map/?msclkid=9e415084b5bb11ec86c6f367b7f93e6c>  
  + accessed: April 6, 2022

<describe the origin of all data in the repository, including data collected from outside sources and new data generated by the investigator(s). If data was accessed from an outside database, the date(s) of data access should also be included.>


## Folder structure, file formats, and naming conventions 

<describe the folders contained in the repository, including what type of files they contain>

<describe the formats of files for the various purposes contained in the repository>

<describe your file naming conventions>

## Metadata

```{r metadata info, echo = FALSE, message=FALSE, warning=FALSE}
countyhealth <- read.csv("../Jacob_Chan_Final/Data/Raw/County Health Data 2021.csv", stringsAsFactors = TRUE)

colnames(countyhealth)

```


<For each data file in the repository, describe the data contained in each column. Include the column name, a description of the information, the class of data, and any units associated with the data. Create a list or table for each data file.> 

## Scripts and code

<list any software scripts/code contained in the repository and a description of their purpose.>

## Quality assurance/quality control

<describe any relevant QA/QC procedures taken with your data. Some ideas can be found here:>
<https://www.dataone.org/best-practices/develop-quality-assurance-and-quality-control-plan>
<https://www.dataone.org/best-practices/ensure-basic-quality-control>
<https://www.dataone.org/best-practices/communicate-data-quality>
<https://www.dataone.org/best-practices/identify-outliers>
<https://www.dataone.org/best-practices/identify-values-are-estimated>


