---
title: "ChanJacob_README"
author: "Meilin Chan and Natasha Jacob"
date: "4/6/2022"
output: pdf_document
---

# <ChanJacob_ENV872_EDA_FinalProject>
```{r setup, include = FALSE, message = FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(dplyr)

```

<Instructions: copy and paste this template into your project README file (found in the parent folder of the repository). Fill in relevant information as requested.>

<General notes: add as much information as is relevant for your repository. Some overarching guidelines are provided, but feel free to expand on these guidelines.>
<More resources found here: https://www.dataone.org/all-best-practices>
<Delete the text inside the brackets when formatting your file.>

## Summary

> The data within this repository includes a shapefile to conduct any spatial analysis, a raw data file which includes public health and population distribution data per NC county, and a data file which includes the spatial locations of hog farms within NC. The purpose of this repository is to examine if there are higher levels of health challenges and disparities within NC with high concentrations of hog farms compared to other NC counties, as well as observe any racial and economic inequities within communities where hog farms are mainly located. We will compare racial, health, etc. distributions within North Carolina compared to counties with the highest concentration of hog farms. 


## Investigators

**Meilin Chan**  
*meilin.chan@duke.edu*  
  
**Natasha Jacob**  
*natasha.jacob@duke.edu*  


## Keywords

* hog farms
* environmental racism
* environmental health
* health disparities

## Database Information

* County Health Data 2021.xlsx (Raw folder)
  + Source: [NCIOM Data] <https://nciom.org/nc-health-data/map/?msclkid=9e415084b5bb11ec86c6f367b7f93e6c>  
  + accessed: April 6, 2022
  
* County Health Data 2021.csv (Process folder)
  + Source: wrangled County Health Data 2021.xlsx to csv format
  
* cb_2019_us_county_20m.shp (Raw folder)
  + Source: Environmental_Data_Analytics_2022 Class GitHub Repository (Folder: Data/Spatial)


## Folder structure, file formats, and naming conventions 
+ Folders
  *Data = stores all data used in the project
    * Raw = raw data, unwrangled
    * Processed = wrangled data
  * README = folder containing the template and project-specific version of README which provides background and overview of the repository and purpose of the project.

+ Types of files
  *.csv = for wrangling and creation of dataframes for ggplot visualizations and other analyses
  *.shp = for spatial analyses
  
+ Naming of files
  * Raw data file names are kept as the original file names when pulled from the source data bases. 
  * Processed data file names include a short description of what the files contain


## Metadata

### Raw Folder
> We wrangled the original County Health Data 2021.xlsx into County Health Data 2021.csv file for easier access with manipulating the dataset.The County Health Data 2021.csv file has 62 rows and 109 columns. The class of all the columns is a factor. The dataframe contains descriptions for each row (Bold.Indicator.Name.Text) in the Indicator.Descriptor.Text and Source.Label columns. 

```{r raw info, echo = FALSE, message = FALSE, warning = FALSE}

countyraw <- read.csv("../Data/Raw/County Health Data 2021.csv", stringsAsFactors = TRUE)

summary(countyraw)
dim(countyraw)
class(countyraw)

```


### Processed
> The projectNCcounties_HealthData.csv file is our wrangled dataframe in which we conducted analysis on. We wrangled all the columns with filtered into our processed dataset into the numerical class. 

```{r processed info, echo = FALSE, message=FALSE, warning=FALSE}
countyhealth <- read.csv("../Data/Processed/projectNCcounties_HealthData.csv", stringsAsFactors = TRUE)

summary(countyhealth)

dim(countyhealth)

class(countyhealth)

```

|             Column Name               |     Description - all per county     |
|:-------------------------------------:|:------------------------------------:|
|    Population                         |   Total population                   |
|    Child_Population_percentage        |   % (per county)                     |
|    Elderly_Population_percentage      |   % (per county)                     |
|    White_Population_percentage        |   % (per county)                     |
|    Hispanic_Latinx_percentage         |   % (per county)                     |
|    African_American_percentage        |   % (per county)                     |
|    American_Indian_percentage         |   % (per county)                     |
|    Reading_Proficiency_percentage     |   % literacy rate per county         |
|    Uninsured_Adults_percentage        |   % adults w/out health insurance    |
|    Medicaid_CHIP_enrolles_percentage  |   % individ. w/ Medicaid or CHIP     |
|    Primary_Care_Physicians            |   Physicians per 10,000 pop.         |
|    Low_Birthweight_percentage         |   % of low birth weight occurrences  |
|    Infant_Mortality_Rate              |   % rate of infant mortality         |
|    Cancer_Incidence                   |   Incidence rates per 100,000 pop.   |
|    Heart_Disease                      |   Death rates per 100,000 pop.       |
|    Poverty_percentage                 |   % of pop. living under poverty line|
|    Air_Pollution                      |   avg. daily density of PM (ug/m3)   |

### Spatial
> 

<For each data file in the repository, describe the data contained in each column. Include the column name, a description of the information, the class of data, and any units associated with the data. Create a list or table for each data file.> 

## Scripts and code
> No scripts or code are contained in our repository.

## Quality assurance/quality control
> We sourced our data from reputible sources such as Google Maps, the Environmental_Data_Analytics_2022 GitHub, and the North Carolina Institute of Medicine. To ensure high quality of consistancy in our data, we kept organized unwrangled data in a Raw folder, dataframes we had wrangled and saved in a Processed folder, and any data used with spatial analysis into a Spatial folder. We renamed our data files using names that would provide users a general idea of what the data contained. We never wrangled or edited the Raw data files, and agreed to only save wrangled data that would be used widely in our analyses. All our data and project files were saved and backed up to our GitHub repository which we pulled from at the beginning of each session, and pushed to at the end of each session. 




